Consistent Polyhedral Surrogates via Embedding

### 2019-05-17 neurips-19/poly-embed.tex
new proof of polyhedral implies finite trim; notes on bo-separated

\begin{proof}[Proof of Theorem~\ref{thm:polyhedral-embed}]
  Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss.
  For all $p$, let $P(p)$ be the epigraph of the convex function $u\mapsto \inprod{p}{L(u)}$.
  From Lemma~\ref{lem:polyhedral-pd-same}, we have that the power diagram induced by the projection of $P(p)$ onto $\reals^d$ is the same for any $p\in\inter\simplex$.
  Let $q\in\inter\simplex$ be the uniform distribution on $\Y$, and $V_\Y$ be the set of vertices of $P(q)$ projected onto $\reals^d$.
  By the above, this set is the same had we replaced $q$ by any $p\in\inter\simplex$.
  \raf{I just realized that the focus on vertices is ill-fated: the hinge-with-extra-dimenson example shows that there might not be \emph{any} vertices.  We need to stick to faces (not necessarily facets) of the $P(q)$.}

  Now let $\Gamma := \Gamma[L]$.
  We claim for all $p\in\inter\simplex$, that $\Gamma(p) \cap V_\Y \neq \emptyset$.
  To see this, let $u \in \Gamma(p)$, and $u' = (u,\inprod{p}{L(u)}) \in P(p)$.
  The optimality of $u$ is equivalent to $u$ being contained in the face exposed by the normal $(0,\ldots,0,-1)\in\reals^{d+1}$, which is a face of $P(p)$.
  Let $v'\in\reals^{d+1}$ be a vertex on such a face, and $v\in V_\Y$ its projection onto $\reals^d$.
  Then $v$ is also optimal, and therefore $v\in\Gamma(p)$.

  Now consider $\Y'\subset \Y$.
  Applying the above argument on distributions $p$ with support exactly $\Y'$, we have a similar guarantee: a finite set $V_{\Y'}$ such that $\Gamma(p) \cap V_{\Y'} \neq \emptyset$ for all $p$ with support exactly $\Y'$.
  (When $\Y' = \{y\}$ is a singleton, we simply take the projected vertices of $L(\cdot)_y$.)

  Thus, taking $V = \bigcup_{\Y'\subseteq\Y} V_{\Y'}$, we have for all $p\in\simplex$ that $\Gamma(p) \cap V \neq \emptyset$.
  This implies that $\trim(\Gamma) \subseteq \{\Gamma_v : v\in V\}$, which is finite, so Proposition~\ref{prop:embed-trim} now gives the conclusion.
  \jessiet{Do we need to show the preimage of $\Gamma$ is $V$?}

>>>>

\begin{proof}[Proof of Theorem~\ref{thm:polyhedral-embed}]
  Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss.
  For all $p$, let $P(p)$ be the epigraph of the convex function $u\mapsto \inprod{p}{L(u)}$.
  From Lemma~\ref{lem:polyhedral-pd-same}, we have that the power diagram $D_\Y$ induced by the projection of $P(p)$ onto $\reals^d$ is the same for any $p\in\inter\simplex$.
  Let $\F_\Y$ be the set of faces of $D_\Y$, which by the above are the set of faces of $P(p)$ projected onto $\reals^d$ for any $p\in\inter\simplex$.

  Now let $\Gamma := \Gamma[L]$.
  We claim for all $p\in\inter\simplex$, that $\Gamma(p) \in \F_\Y$.
  To see this, let $u \in \Gamma(p)$, and $u' = (u,\inprod{p}{L(u)}) \in P(p)$.
  The optimality of $u$ is equivalent to $u'$ being contained in the face $F$ of $P(p)$ exposed by the normal $(0,\ldots,0,-1)\in\reals^{d+1}$.
  Thus, $\Gamma(p) = \argmin_{u\in\reals^d} \inprod{p}{L(u)}$ is a projection of $F$ onto $\reals^d$, which is an element of $\F_\Y$.

  % Let $v'\in\reals^{d+1}$ be a vertex on such a face, and $v\in \F_\Y$ its projection onto $\reals^d$.
  % Then $v$ is also optimal, and therefore $v\in\Gamma(p)$.

  Now consider $\Y'\subset \Y$, $\Y'\neq\emptyset$.
  Applying the above argument, we have a similar guarantee: a finite set $\F_{\Y'}$ such that $\Gamma(p) \in \F_{\Y'}$ for all $p$ with support exactly $\Y'$.
  % (When $\Y' = \{y\}$ is a singleton, we simply take the faces of $D_{\{y\}}projected vertices of $L(\cdot)_y$.)

  Thus, taking $\F = \bigcup\{\F_{\Y'} | \Y'\subseteq\Y, \Y'\neq\emptyset\}$, we have for all $p\in\simplex$ that $\Gamma(p) \in \F$.
  This implies that $\trim(\Gamma) \subseteq \F$, which is finite, so Proposition~\ref{prop:embed-trim} now gives the conclusion.
  \jessiet{Do we need to show the preimage of $\Gamma$ is $\F$?}
  \raft{This question doesn't type check... yeah, can't figure out what you are asking.  Maybe that $\Gamma$ is onto $\F$?  We don't need to show that (and it's not true).}

### 2019-05-17 notes/bo-separated.tex
Added two questions/comments to Bo's notes
new proof of polyhedral implies finite trim; notes on bo-separated

\begin{lemma} \label{lemma:U-convex-closed}
  $\mathcal{U}$ is a finite set, and each element $U$ is a closed, convex polyhedron.
\end{lemma}
\begin{proof}
  \bo{More of a sketch right now}
  Let $p$ be in the interior of the simplex.
  For any such $p$, the polyhedral expected loss $L(\cdot;p)$ projects down to the same fixed power diagram over $\reals^d$.
  The minimizer of a polyhedral loss is always the projection of some face of the polyhedron, \bo{cite} so there are a finite number of possible minimizers and each is a closed, convex polyhedron (being a member of the power diagram or an intersection of members).
  Now for $p$ not in the interior, consider each of finitely many subsets of the outcome space and restrict $p$ to have support on this subset; repeat the above arguments.
  The total number of report sets $U$ is still finite, each a closed, convex polyhedron.
\end{proof}

\begin{lemma} \label{lemma:U-surject}
  There is a surjection from $\mathcal{U}$ to $\mathcal{S}$ where $U \mapsto R$ if, for some $p$, we have $\gamma(p) = U$ and $\Gamma(p) = R$.
  We write $R_U$ for the value of this map at $U$.
\end{lemma}
\begin{proof}
  For each report $r$, write $u_r$ for its embedding point and for each $U \in \mathcal{U}$, let $R_U = \{r : u_r \in U\}$.
  Recall that $\mathcal{U}$ is the collection of sets $U = \gamma(p)$ for some $p$, while $\mathcal{S}$ is the collection of sets $R = \Gamma(p)$ for some $p$.
  By definition of embedding, we have $u_r \in \gamma(p) \iff r \in \Gamma(p)$.
  So if $U = \gamma(p)$, we must have $R = \Gamma(p)$.
  This shows that $R_U \in \mathcal{S}$ for all $U$, and furthermore, the map is surjective because every $R \in \mathcal{S}$ is equal to $R_U$ for some $U \in \mathcal{U}$.
\end{proof}

\begin{lemma} \label{lemma:calibrated-pos}
  If a subset of $\mathcal{U}$ intersect, then their corresponding report sets intersect as well.
\end{lemma}
\begin{proof}
  \bo{already have this: it is our proof that a calibrated link is possible!}
\end{proof}

@@

\begin{lemma} \label{lemma:thick-nonempty}
  Let $\{U_j : j \in \mathcal{J}\}$ be a finite collection of closed, convex sets with nonempty joint intersection.
  Then there exists  $\epsilon > 0$ such that $\cap_j B(U_j,\epsilon) \subseteq B(\cap_j U_j, \delta)$.
\end{lemma}
\begin{proof}
  By induction on the size of the collection.
  If the size is $1$, set $\epsilon = \delta$.
  If the size is larger than $1$, let $U_j$ be any member of the collection, let $U' = \cap_{j'\neq j} U_{j'}$, and let $C(\epsilon) = \cap_{j' \neq j} B(U_{j'},\epsilon)$.
  Let $D = U_j \cap U'$.
  We must show that $B(U_j,\epsilon) \cap C(\epsilon) \subseteq B(D,\delta)$.
  By Lemma \ref{lemma:enclose-halfspaces}, we can enclose $D$ strictly within a polyhedron $D'$, the intersection of a finite number of open halfspaces, which is itself strictly enclosed in $B(D,\delta)$.
  (For example, if $D$ is a point, then enclose it in a hypercube, which is enclosed in the ball $B(D,\delta)$.)
%  \raf{We can use the fact that the $U_j$ are all polyhedral here, so $D$ is polyhedral as well.  Now let $B_1()$ be a ball with respect to $\|\cdot\|_1$, and observe that $B_1(D,\delta')$ is therefore also polyhedral for all $\delta'$.  Using norm bounds, we can find a small enough $\delta'$ so that $B_1(D,\delta') \subseteq B(D,\delta)$.}
  We will prove that, for small enough $\epsilon$, $B(U_j,\epsilon) \cap C(\epsilon)$ is contained in $D'$.
  This implies that it is contained in $B(D,\delta)$.

  For each halfspace defining $D'$, consider its complement $F$, a closed halfspace.
  We prove that $F \cap B(U_j,\epsilon) \cap C(\epsilon) = \emptyset$.
  Consider the intersections of $F$ with $U$ and $U'$, call them $G$ and $G'$.
  These are closed, convex sets that do not intersect (because $D$ in contained in the complement of $F$).
  So $G$ and $G'$ are separated by a nonzero distance, so $B(G,\gamma) \cap B(G',\gamma) = \emptyset$ for small enough $\gamma$.
  And $B(G,\gamma) = F \cap B(U_j,\gamma)$ while $B(G',\gamma) = F \cap B(U',\gamma)$.
  This proves that $F \cap B(U_j,\gamma) \cap B(U',\gamma) = \emptyset$.
  By inductive assumption, $C(\epsilon) \subseteq B(U',\gamma)$ for small enough $\epsilon$.
  So $F \cap B(U_j,\gamma) \cap C(\epsilon) = \emptyset$.

  We now take the minimum over these finitely many $\epsilon$'s (one per halfspace).
\end{proof}

@@

\begin{lemma} \label{lemma:thick-empty}
  Let $\{U_j : j \in \mathcal{J}\}$ be a finite collection of nonempty closed, convex sets with empty joint intersection.
  Then for all $\delta > 0$, there exists  $\epsilon > 0$ such that $\cap_j B(U_j,\epsilon) = \emptyset$.
\end{lemma}
\begin{proof}
  By induction on the size of the family.
  Note that the family must have size at least two.
  Let $U_j$ be any set in the family and let $U' = \cap_{j' \neq j} U_{j'}$.
  There are two possibilities.

@@

\begin{corollary} \label{cor:thick-intersect}
  There exists a small enough $\epsilon > 0$ such that, for any subset $\{U_j : j \in \mathcal{J}\}$ of $\mathcal{U}$, if $\cap_j U_j = \emptyset$, then $\cap_j B(U_j,\epsilon) = \emptyset$.
\end{corollary}
\begin{proof}
  For each subset, Lemma \ref{lemma:thick-empty} gives an $\epsilon$.
  We take the minimum over these finitely many choices.
\end{proof}

@@

>>>>

\begin{lemma} \label{lemma:U-convex-closed}
  $\mathcal{U}$ is a finite set, and each element $U$ is a closed, convex polyhedron.
\end{lemma}
\begin{proof}
  \raf{I updated the main doc to include a proof of exactly this statement.  It made me realize that the existing proof was wrong, and this is actually the right thing to show anyway!}
  \bo{More of a sketch right now}
  Let $p$ be in the interior of the simplex.
  For any such $p$, the polyhedral expected loss $L(\cdot;p)$ projects down to the same fixed power diagram over $\reals^d$.
  The minimizer of a polyhedral loss is always the projection of some face of the polyhedron, \bo{cite} so there are a finite number of possible minimizers and each is a closed, convex polyhedron (being a member of the power diagram or an intersection of members).
  Now for $p$ not in the interior, consider each of finitely many subsets of the outcome space and restrict $p$ to have support on this subset; repeat the above arguments.
  The total number of report sets $U$ is still finite, each a closed, convex polyhedron.
\end{proof}

\begin{lemma} \label{lemma:U-surject}
  There is a surjection from $\mathcal{U}$ to $\mathcal{S}$ where $U \mapsto R$ if, for some $p$, we have $\gamma(p) = U$ and $\Gamma(p) = R$.
  We write $R_U$ for the value of this map at $U$.
\end{lemma}
\begin{proof}
  For each report $r$, write $u_r$ for its embedding point and for each $U \in \mathcal{U}$, let $R_U = \{r : u_r \in U\}$.
  Recall that $\mathcal{U}$ is the collection of sets $U = \gamma(p)$ for some $p$, while $\mathcal{S}$ is the collection of sets $R = \Gamma(p)$ for some $p$.
  \jessiet{Doesn't $U$ correspond to $\Gamma$ and $R$ to $\gamma$?}
  By definition of embedding, we have $u_r \in \gamma(p) \iff r \in \Gamma(p)$.
  So if $U = \gamma(p)$, we must have $R = \Gamma(p)$.
  This shows that $R_U \in \mathcal{S}$ for all $U$, and furthermore, the map is surjective because every $R \in \mathcal{S}$ is equal to $R_U$ for some $U \in \mathcal{U}$.
\end{proof}

The next lemma shows that if a subset of $\mathcal{U}$ intersect, then their corresponding report sets intersect as well.
\begin{lemma} \label{lemma:calibrated-pos}
  Let $\mathcal{U}' \subseteq \mathcal{U}$.
  If $\cap_{U\in\mathcal{U}'} U \neq \emptyset$ then $\cap_{U\in\mathcal{U}'} R_U \neq \emptyset$.
\end{lemma}
\begin{proof}
  \bo{already have this: it is our proof that a calibrated link is possible!}
\end{proof}

@@

\begin{lemma} \label{lemma:thick-nonempty}
  Let $\{U_j : j \in \mathcal{J}\}$ be a finite collection of closed, convex sets with $\cap_{j\in\mathcal{J}} U_j \neq \emptyset$.
  Then there exists  $\epsilon > 0$ such that $\cap_j B(U_j,\epsilon) \subseteq B(\cap_j U_j, \delta)$.
  \raft{FYI, formalized some of the English in both the statement and proof}
\end{lemma}
\begin{proof}
  We induct on $|\mathcal{J}|$.
  If $|\mathcal{J}|=1$, set $\epsilon = \delta$.
  If $|\mathcal{J}|>1$, let $j\in\mathcal{J}$ be arbitrary, let $U' = \cap_{j'\neq j} U_{j'}$, and let $C(\epsilon) = \cap_{j' \neq j} B(U_{j'},\epsilon)$.
  Let $D = U_j \cap U'$.
  We must show that $B(U_j,\epsilon) \cap C(\epsilon) \subseteq B(D,\delta)$.
  By Lemma \ref{lemma:enclose-halfspaces}, we can enclose $D$ strictly within a polyhedron $D'$, the intersection of a finite number of open halfspaces, which is itself strictly enclosed in $B(D,\delta)$.
  (For example, if $D$ is a point, then enclose it in a hypercube, which is enclosed in the ball $B(D,\delta)$.)
%  \raf{We can use the fact that the $U_j$ are all polyhedral here, so $D$ is polyhedral as well.  Now let $B_1()$ be a ball with respect to $\|\cdot\|_1$, and observe that $B_1(D,\delta')$ is therefore also polyhedral for all $\delta'$.  Using norm bounds, we can find a small enough $\delta'$ so that $B_1(D,\delta') \subseteq B(D,\delta)$.}
  We will prove that, for small enough $\epsilon$, $B(U_j,\epsilon) \cap C(\epsilon)$ is contained in $D'$.
  This implies that it is contained in $B(D,\delta)$.

  For each halfspace defining $D'$, consider its complement $F$, a closed halfspace.
  We prove that $F \cap B(U_j,\epsilon) \cap C(\epsilon) = \emptyset$.
  Consider the intersections of $F$ with $U$ and $U'$, call them $G$ and $G'$.
  These are closed, convex sets that do not intersect (because $D$ in contained in the complement of $F$).
  So $G$ and $G'$ are separated by a nonzero distance, so $B(G,\gamma) \cap B(G',\gamma) = \emptyset$ for small enough $\gamma$.
  And $B(G,\gamma) = F \cap B(U_j,\gamma)$ while $B(G',\gamma) = F \cap B(U',\gamma)$.
  This proves that $F \cap B(U_j,\gamma) \cap B(U',\gamma) = \emptyset$.
  By inductive assumption, $C(\epsilon) \subseteq B(U',\gamma)$ for small enough $\epsilon = \epsilon_F$.
  So $F \cap B(U_j,\gamma) \cap C(\epsilon) = \emptyset$.
  We now let $\epsilon$ be the minimum over these finitely many $\epsilon_F$ (one per halfspace).
\end{proof}

@@

\begin{lemma} \label{lemma:thick-empty}
  Let $\{U_j : j \in \mathcal{J}\}$ be a finite collection of nonempty closed, convex sets with $\cap_{j\in\mathcal{J}} U_j = \emptyset$.
  Then for all $\delta > 0$, there exists  $\epsilon > 0$ such that $\cap_{j\in\mathcal{J}} B(U_j,\epsilon) = \emptyset$.
\end{lemma}
\begin{proof}
  By induction on the size of the family.
  Note that the family must have size at least two.
  Let $U_j$ be any set in the family and let $U' = \cap_{j' \neq j} U_{j'}$.
  There are two possibilities.

@@

\begin{corollary} \label{cor:thick-intersect}
  There exists a small enough $\epsilon > 0$ such that, for any subset $\{U_j : j \in \mathcal{J}\}$ of $\mathcal{U}$, if $\cap_j U_j = \emptyset$, then $\cap_j B(U_j,\epsilon) = \emptyset$.
\end{corollary}
\jessiet{Mention $\mathcal{U}$ is finite for good measure?}
\raft{I think the proof makes it clear enough}
\begin{proof}
  For each subset, Lemma \ref{lemma:thick-empty} gives an $\epsilon$.
  We take the minimum over these finitely many choices.
\end{proof}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
